---
title: "Cat Observations Keyword Search"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_collapsed: true
    toc_depth: 2
    number_sections: false
vignette: >
  %\VignetteIndexEntry{cat-search-template}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Introduction

This vignette will walk you through how to run a comprehensive search of the `cat_observations` data set. The search allows for "fuzzy matching" of keywords by utilizing regex. **_Pro Tip:_** Test your custom regex here: [https://regex101.com/](https://regex101.com/)

This tutorial uses several custom `catscan` functions to create regex search codes, search those words within your data set, and produce a nicely formatted xlsx report. Although these functions were built with the cat observation data in mind, they can be generalized to any data set search. 

# Example search 

This vignette will walk through using the `catscan` functions to provide a search report for the following request:

> **Media Request**
> 
> Email from Jenny Rieck on 2026-04-01
>
> Please provide all cat incidents for Mr. Jones, aka "MJ" since Feb 01 2026. Please indicate if the incident note mentions that he was being noisy, meowing, or yowling. 

## Load the data

First, load the data set with the text you wish to search (in this case `cat_observations`) and filter for in-scope incidents based on the timeframe of interest.

```{r load_data}
# Load packages
library(catscan)
library(dplyr)
library(tidyr)
library(openxlsx)

# Report Name
search_name = "Cat Search - Mr Jones"

# In-scope criteria
start_date = "2026-02-01"
end_date = Sys.Date()

# Load Data and filter date range
data("cat_observations")
cat_observations <- cat_observations |>
  filter(date_received >= start_date,
         date_received <= end_date)
```

## Craft search terms

Now it's time to craft your list of keywords to search in your data set. All search terms are treated as `OR` conditions, meaning at least one of them must be present to constitute a "hit" for a particular case.

Some basic regex formatting is handled by input parameters to the `mk_keyword_search_regex()` function in `catscan`:

* `STANDALONE` - TRUE (default) to set word boundaries (i.e., stand alone words) or FALSE to search the keyword within a word
* `IGNORE_CASE` - TRUE (default) to ignore upper/lower case
* `MAKE_PLURAL` - FLASE (default) or TRUE to make a keyword plural by adding an optional "s" to the end 

In addition to the simple regex provided by `mk_keyword_search_keycode()`, you can include regex within your search terms. For example, `"nois(y|e)"` will hit for both "noisy" and "noise". 

```{r search_terms}
# Noisy Mr Jones search terms:
search_terms <- c("(Mister|Mr\\.?) Jones?", "meow", "yowl", "nois(y|e)")

search_keycodes <- mk_keyword_search_regex(search_terms, 
                                             STANDALONE = FALSE,
                                             IGNORE_CASE = TRUE,
                                             MAKE_PLURAL = FALSE)
```

## Run the search

The `keyword_search()` function in `catscan` will take the data frame created by `search_keycodes()`, search your given data for those terms, and return only those rows where a “hit” was found. You will need to specify the name of the  column with the text you wish to search with the `SEARCH_COLUMN` parameter. In this case, we are searching the `observation_note` column for our keywords generated above.

```{r run_seach}
search_results <- keyword_search(
  KEYWORDS_DF = search_keycodes, 
  DATA = cat_observations, 
  SEARCH_COLUMN = "observation_note"
) 
```

## Aggregate the results

Because `keyword_search()` returns one row for each hit, text fields where multiple keywords are found will be duplicated. Therefore, you may want to reshape the search results to aggregate all the keyword hits across each observation.

Because we are only interested in observations that involved Mr. Jones, we will further filter the search results for only those hits that include those terms (Mr. Jones or MJ). We will also add a column for the fiscal year that the incident was received.

```{r aggregate}
search_results_aggregated <- search_results |>
  group_by(note_id) |>
  mutate(ALL_HITS = paste0(unique(MATCH), collapse = " | ")) |>
  ungroup() |>
  select(-MATCH) |>
  distinct() |>
  filter(when_any(
    grepl("Mister", ALL_HITS),
    grepl("MJ", ALL_HITS),
  )) |>
  mutate(fy_received = gc_fiscal_year(date_received)) |>
  arrange(incident_id)
```

## Save the report

Finally, `catscan` has several functions to format an standard excel report with two tabs. The first tab, created with `wb_search_params()`, includes the search parameters for the timeframe, fields, and search terms. The second tab, created with `wb_summary_tab()` includes the search results table created in the step above. Custom Excel styling is also applied with `wb_style()`.

```{r report_out}
# Package report parameters for the wb_ helpers
wb_params <- list(
  title = search_name, 
  start_date = start_date, 
  end_date = end_date, 
  search_terms = data.frame(Search_Terms = search_terms),
  search_fields = data.frame(Search_Fields = "Observation Notes")
)

# Create an empty openxlsx workbook
wb <- createWorkbook()

# Add the search parameters
addWorksheet(wb, sheetName = "Search Parameters", gridLines = FALSE)
wb_search_params(wb = wb, 
                 wb_style = wb_style(), 
                 params = wb_params, 
                 sheet_num = 1)

# Add the search results
addWorksheet(wb, sheetName = "Search Results", gridLines = FALSE)

# Specify column names to wrap or widen
to_wrap <- c("observation_note")
to_widen <- c("cat_name", "observation_note", "ALL_HITS")

wb_summary_tab(wb = wb,
               wb_style = wb_style(),
               summary_tab = search_results_aggregated,
               sheet_num = 2,
               cols_to_wrap = to_wrap,
               cols_to_widen = to_widen)

# Save the xlsx
saveWorkbook(wb, paste0(wb_params$title, "_" , Sys.Date(), ".xlsx"), overwrite = TRUE)
```


# Full search template

A full script template for conducting a keyword search can be found below:

```{r full_template}
####
## Request ----
####
## Email from Jenny Rieck on 2026-04-01
## Media Request - Please provide all cat incidents for Mr. Jones, aka MJ since 
## Feb 01 2026. Please indicate if the incident note mentions that he was
## noisy, meowing, or yowling

library(catscan)
library(dplyr)
library(tidyr)
library(openxlsx)

####
## SETUP ----
####

# Specify search parameters
search_params <- list(
  title = "Cat Search - Mr Jones",
  start_date = "2026-02-01",
  end_date = Sys.Date(),
  search_terms = data.frame(
    Search_Terms = c("(Mister|Mr\\.?) Jones?", "meow", "yowl", "nois(y|e)")
  ),
  search_fields = data.frame(
    Search_Fields = "Observation Notes"
  ) 
)

# Load Data and filter date range
data("cat_observations")
cat_observations <- cat_observations |>
  filter(date_received >= search_params$start_date,
         date_received <= search_params$end_date)

####
## SEARCH ----
####

# Create keyword regex
search_keycodes <- mk_keyword_search_regex(
  search_params$search_terms$Search_Terms, 
  STANDALONE = FALSE,
  IGNORE_CASE = TRUE,
  MAKE_PLURAL = FALSE
)

# Run the search:
search_results <- keyword_search(
  KEYWORDS_DF = search_keycodes, 
  DATA = cat_observations, 
  SEARCH_COLUMN = "observation_note"
)

# Aggregate hits across notes:
search_results_aggregated <- search_results |>
  group_by(note_id) |>
  mutate(ALL_HITS = paste0(unique(MATCH), collapse = " | ")) |>
  ungroup() |>
  select(-MATCH) |>
  distinct() |>
  filter(when_any(
    grepl("Mister", ALL_HITS),
    grepl("MJ", ALL_HITS),
  )) |>
  mutate(fy_received = gc_fiscal_year(date_received)) |>
  arrange(incident_id)

####
## Report Out ----
####

# Create an empty openxlsx workbook
wb <- createWorkbook()

# Add the search parameters
addWorksheet(wb, sheetName = "Search Parameters", gridLines = FALSE)
wb_search_params(wb = wb, 
                 wb_style = wb_style(), 
                 params = search_params, 
                 sheet_num = 1)

# Add the search results
addWorksheet(wb, sheetName = "Search Results", gridLines = FALSE)

to_wrap <- c("observation_note")
to_widen <- c("cat_name", "observation_note", "ALL_HITS")

wb_summary_tab(wb = wb,
               wb_style = wb_style(),
               summary_tab = search_results_aggregated,
               sheet_num = 2,
               cols_to_wrap = to_wrap,
               cols_to_widen = to_widen)

# Save the xlsx
saveWorkbook(wb, paste0(search_params$title, "_" , Sys.Date(), ".xlsx"), overwrite = TRUE)
```
